---
title: "UT_B117_main"
output: html_document
---
This script loads in the raw data files and calls the functions that run the logistic growth model of B117 vs wildtype and the SEIR model the projects infections
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(EpiEstim)
library(tidyr)
library(EnvStats)
library(reshape2)
library(cowplot)
library(knitr)
library(flexmix)
library(zoo)
library(DescTools)
library(dplyr)
source("UT_B117_functions.R")
theme_set(theme_cowplot())
ggplot + background_grid(major = "xy", minor = "xy")
options(warn = -1)
```
Load in SGTF data and PCT case data
```{r}
DATA_PATH_CTS<-'../data/pctData_posOnly_210505.csv'
DATA_PATH_CASES<-'../data/UT_data_streams.csv'
DATA_PATH_INTROS<-'../data/introductions_Jan15.csv'

```
Dates and parameters for analysis
```{r}
spring_arrival_date<-as.Date("2021-01-16", format = "%Y-%m-%d")
spring_last_date<-as.Date("2021-04-09", format = "%Y-%m-%d")
spring_end_semester_date<-as.Date("2021-05-23", format = "%Y-%m-%d")
initial_estimate_date<-as.Date("2021-02-12", format = "%Y-%m-%d")
fall_arrival_date<-as.Date("2020-08-20", format = "%Y-%m-%d")
fall_last_date<-as.Date("2020-12-19", format = "%Y-%m-%d")

# Some global parameters
QC_thres<-28
delta<-1/7 #(1/duration of infectiousness)
gamma<- 1/3 # (1/duration of latent period)
dt<-0.1
mean_GI<-5.8 # average generation interval (for EpiEstim)
std_GI<-2.3 # std dev generation interval (for EpiEstim)
N_POPULATION <-30000 # estimated "active" population at UT
prop_prev_inf_Aug = 0.05 
nsamps<-500 # number of samples from the distributions for uncertainty
```
Process the Ct data to get SGTF and non-SGTF positives from the Ct values of the 3 target genes
```{r}
Ct_data<-read.csv(DATA_PATH_CTS) %>%
  mutate(test_date = as.Date(test_date))
Ct_data<-Ct_data %>%select(test_date, N.gene.Ct, S.gene.Ct, ORF1ab.Ct)%>%
  filter(test_date>= spring_arrival_date)%>%
  filter(test_date<=spring_last_date)
Ct_data<-Ct_data[order(Ct_data$test_date),]
Ct_data<-Ct_data%>%rowwise()%>%mutate(mean_Ct = mean(c(N.gene.Ct, S.gene.Ct, ORF1ab.Ct), na.rm = TRUE))
Ct_data['S_dropout']<-is.na(Ct_data$S.gene.Ct)
Ct_data$poss_B117<-Ct_data$S_dropout==TRUE & Ct_data$N.gene.Ct<QC_thres
Ct_data['strain']<-rep(NA, nrow(Ct_data))
for (i in 1:nrow(Ct_data)){
  if (Ct_data$S_dropout[i] == TRUE){
    Ct_data$strain[i]<-'SGTF'
  }
  else 
    Ct_data$strain[i]<-'non_SGTF'
}

#Make a day to day summary of SGTF presence
Ct_data_daily_long<-Ct_data%>%filter(mean_Ct<QC_thres)%>%
  group_by(test_date, strain)%>%summarise(n = n())
Ct_data_daily_summary<-Ct_data_daily_long%>%
  pivot_wider(names_from = strain, values_from = n, values_fill = 0)
Ct_data_daily_summary<-Ct_data_daily_summary%>%group_by(test_date)%>%
  mutate(n_Sdrop = SGTF,
         n_Spos = non_SGTF,
         n_pos = SGTF+non_SGTF,
         prev_SGTF = SGTF/n_pos,
         SE_prev = sqrt(prev_SGTF*(1-prev_SGTF)/n_pos))
# Export the daily data for fitting. Will use this to feed into Woody's SGTF.R code
write.csv(Ct_data_daily_summary, '../data/Ct_data.csv')
# Make a dataset used for the initial fitting from data through Feb.
Ct_data_init<-Ct_data_daily_summary%>%filter(test_date<=initial_estimate_date)
write.csv(Ct_data_init, '../data/Ct_data_init.csv', row.names = FALSE)


# Make figures with frequency of SGTF over time as well as cases colored by SGTF status
path ="../out/dailySGTF.png"
plt_freq<-ggplot(Ct_data_daily_summary, aes(x = test_date, y = 100*(prev_SGTF)), color = "black") + geom_point() + geom_errorbar(aes(ymin = 100*(prev_SGTF - SE_prev), ymax = 100*(prev_SGTF+SE_prev)), width = 0.2, position = position_dodge(0.5))+
xlab ('Test Collection Date') + ylab ('Frequency (%) of cases that are S dropouts') + ggtitle('Daily frequency of S dropouts with low Ct values')
plt_freq
ggsave(filename = path, device = "png", dpi = 700) 

path = "../out/stackedbarSGTF.png"
stacked<-ggplot(Ct_data_daily_long, aes(x = test_date, y = n, fill = strain))  + geom_bar(position = "stack", stat = "identity") + xlab('Test Collection Date') + ylab('Number of low Ct positives') + theme(legend.position="top")+ scale_fill_discrete(name = "S gene presence", labels = c("S+", "S-"))
stacked
ggsave(filename = path, device = "png", dpi = 700)

# Use SGTF data set to call function to fit the logistic model with uncertainty from SGTF presence. 

#output: c and k in logistic growth model (intercept and growth rate), current B117 prevalence
# Early estimate: Jan 16- Feb. 12 (probably won't need this)
c1 = -2.861
c1_distrib<-runif(n= nsamps, min = -2.9, max = -2.83)
k1 = 0.077
k1_distrib<-runif(n= nsamps, min = 0.017, max = 0.140) # logistic growth rate parameter bounds

# Use later estimate (Jan 16- April 9) for p(t) and transmissibiltiy advantage
c = -2.861
c_distrib<-runif(n= nsamps, min = -2.9, max = -2.83)
k_p = 0.040
k_p_distrib = runif(n=nsamps, min = 0.037, max = 0.043)
p_curr = 0.627
p_curr_distrib<-runif(nsamps, 0.501, 0.739)




write.csv(c_distrib, '../data/c_distrib.csv', row.names = FALSE)
write.csv(k_p_distrib,'../data/k_p_distrib.csv', row.names = FALSE)
write.csv(p_curr_distrib, '../data/p_curr_distrib.csv', row.names = FALSE)

M1_distrib = exp(mean_GI*k1_distrib)
M2_distrib = exp(mean_GI*k_p_distrib) # This should probably be a function of b0 and b1? 
M_CI<-quantile(M2_distrib, probs = c(0.025, 0.5, 0.975))

```
Load in the logistic growth model outputs (posterior distributions of the y-intercept and growth rate, and the current estimate of the B117 prevalence)** These will be used as inputs into the SEIR model**
```{r}

k_p_distrib<-read.csv('../data/k_p_distrib.csv')
k_p_distrib<-k_p_distrib$x
c_distrib<-read.csv('../data/c_distrib.csv')
c_distrib<-c_distrib$x
p_curr_distrib<-read.csv('../data/p_curr_distrib.csv')
p_curr_distrib<-p_curr_distrib$x

# From the literature (Davies 2021 US estimate of 59% more transmissibile (56-63))
k_tr_adv_distrib<-rtri(n = nsamps, min = log(1.56)/5.5, max = log(1.63)/5.5, mode = log(1.59)/5.5)
```

Load in the UT case data, clean and estimate Rt
```{r}
case_data<-read.csv(DATA_PATH_CASES)%>%mutate(date = as.Date(CollectionDate, format = "%Y-%m-%d"))%>%
                                                select(date, student_pos, student_PCT_pos,
                                                       total_student_PCT_tests)%>%
                                                filter(date>=fall_arrival_date, date<=spring_last_date+7)
# Add some columns to the case data
case_data<-case_data%>%mutate(positivity = student_PCT_pos/total_student_PCT_tests)
sevendayPCTpos = c(rep(NA,3), rollapply(case_data$positivity, 7, function(x) mean(x, na.rm = TRUE)), rep(NA,3))
sevendayPCTcases = c(rep(NA,3), rollapply(case_data$student_PCT_pos, 7, function(x) mean(x, na.rm = TRUE)), rep(NA,3))
sevendaycases = c(rep(NA,3), rollapply(case_data$student_pos, 7, function(x) mean(x, na.rm = TRUE)), rep(NA,3))
case_data$sevendayPCTpos = sevendayPCTpos
case_data$sevendayPCTcases = sevendayPCTcases
case_data$sevendaycases = sevendaycases

case_data%>%ggplot() + geom_line(aes(x = date, y = sevendayPCTpos)) + geom_bar(aes(x = date, y = positivity), stat = "identity") + ylab('PCT positivity') + xlab('Test Collection Date')

case_data%>%ggplot() + geom_line(aes(x = date, y = sevendayPCTcases)) + geom_bar(aes(x = date, y = student_PCT_pos), stat = "identity") + ylab('Student PCT cases') + xlab('Test Collection Date')

# Estimate R(t) in the fall and in the spring.
#daily_p_local<-get_daily_p_local(case_data, fall_arrival_date)
case_data_fall<-case_data%>%filter(date<=fall_last_date)
case_data_spring<-case_data%>%filter(date>=spring_arrival_date)
cases_fall<-case_data_fall$student_pos # for fall, we use total positives

avg_tests_per_day<-mean(case_data_spring$total_student_PCT_tests, na.rm = TRUE)
case_data_spring$positivity[is.na(case_data_spring$positivity)]=case_data_spring$sevendayPCTpos[is.na(case_data_spring$positivity)]
# Still need to do something to account for 10-day pause during the Texas winter freeze where no tests were collected, use the 3 days before and after it to impute. 
case_data_around_freeze<-case_data%>%filter(date<=as.Date("2021-02-24"), date>=as.Date("2021-02-10"),
                                            total_student_PCT_tests>0)
avg_positivity_freeze<-sum(case_data_around_freeze$student_PCT_pos)/sum(case_data_around_freeze$total_student_PCT_tests)
avg_n_cases_freeze<-sum(case_data_around_freeze$student_pos)/nrow(case_data_around_freeze)
case_data_spring$positivity[is.na(case_data_spring$positivity)]=avg_positivity_freeze
cases_spring<-round(avg_tests_per_day*case_data_spring$positivity,0) # for spring, we used PCT positivity
case_data_spring$imputed_cases<-cases_spring


Rt_summary_fall<-Rt_fxn_cases(cases_fall, case_data_fall, fall_arrival_date, fall_last_date, mean_GI,std_GI)
Rt_summary_spring<-Rt_fxn_cases(cases_spring, case_data_spring, spring_arrival_date, spring_last_date+7, mean_GI,std_GI)

# Export the Rt data for use in the SEIR model
write.csv(Rt_summary_fall, '../data/Rt_students_fall.csv', row.names = FALSE)
write.csv(Rt_summary_spring, '../data/Rt_students_spring.csv', row.names = FALSE)

```
Load in Rt estimates from fall and spring to generate figures 
```{r}
Rt_summary_fall<-read.csv('../data/Rt_students_fall.csv')
Rt_summary_spring<-read.csv('../data/Rt_students_spring.csv')
Rt_summary_fall$dates<-as.Date(Rt_summary_fall$dates)
Rt_summary_spring$dates<-as.Date(Rt_summary_spring$dates)

Rt_summary_fall%>%ggplot() + geom_line(aes(x =dates, y = Rt_medians))+ geom_ribbon(aes(x = dates, ymin = Rt_lowerbounds, ymax = Rt_upperbounds), alpha = 0.1)

Rt_summary_spring%>%ggplot() + geom_line(aes(x =dates, y = Rt_medians))+ geom_ribbon(aes(x = dates, ymin = Rt_lowerbounds, ymax = Rt_upperbounds), alpha = 0.1)


path = "../out/student_cases.png"
cases<-ggplot(case_data_spring) + geom_line(aes(x = date, y = sevendayPCTcases)) + geom_bar(aes(x = date, y = student_PCT_pos), stat = "identity") + xlab('Date') + ylab('Reported PCT cases') 
cases
ggsave(filename = path, device = "png", dpi = 700)

tspring<-seq(from = 0, to = nrow(case_data_spring)-1, by =1)
 p_t<-(1/(1+exp(-(c+k_p*tspring))))
plot(tspring, p_t)


case_data_spring$n_est_B117<-case_data_spring$student_PCT_pos*p_t
case_data_spring$n_est_WT<-case_data_spring$student_PCT_pos*(1-p_t)
case_data_long<-case_data_spring%>%select(date, n_est_B117, n_est_WT)%>%
  pivot_longer(cols = starts_with("n_est_"), names_to = "strain", names_prefix = "n_est_", values_to = "n", values_drop_na = TRUE)

path = "../out/student_cases_by_B117.png"
cases<-ggplot(case_data_long, aes(x = date, y = n, fill = strain)) + geom_bar(position = "stack", stat = "identity") + xlab('Date') + ylab('Reported PCT cases') +scale_fill_brewer(palette = "Set2")
cases
ggsave(filename = path, device = "png", dpi = 700)

```
Run the SEIR model using the fall case data to estimate the number previously infected
```{r}
fall_intros<-read.csv(DATA_PATH_INTROS)%>%filter(Location == "total", Time == "Fall")%>%select(ML_infections_lo, ML_infections_med, ML_infections_hi)
I0bounds<-c(fall_intros$ML_infections_lo, fall_intros$ML_infections_hi)

par_table<-data.frame(dt, delta, gamma, N_POPULATION, prop_prev_inf_Aug, nsamps)

cumIJan<-run_fall_SEIR(par_table, I0bounds, Rt_summary_fall)
cumICI<-(cumIJan/N_POPULATION)%>%quantile(probs = c(0.025, 0.5, 0.975))
print(cumICI)
```
Use the previous infection estimates from the fall, spring Rt, the current estimates of variant prevalence, and variant increase in transmissibility (from both model estimates) to project infections with and without the variant.
```{r}

spring_par_table<-data.frame(dt, delta, gamma, N_POPULATION, nsamps)
iRts<-c(18, 14) # indices in spring Rt that correspond to the faster and slower spread scenarios

# Output from logistic model is used for B117 prevalence timeline and transmissibility advantage
out_list<-run_spring_SEIR(spring_par_table, case_data_spring, Rt_summary_spring, iRts, spring_last_date, cumIJan, k_p_distrib, k_p_distrib, c_distrib, p_curr_distrib)

df_t_sim<-out_list[[1]] # contains the time course of runs for 500 simulations for both scenarios
df_t<-out_list[[2]] # contains the median and bounds of the time course runs for both scenarios
df_summary<-out_list[[3]] # contains all the summary statistics for static variables

# Ru again, but this time use literature for transmissibility advantage
out_list<-run_spring_SEIR(spring_par_table, case_data_spring, Rt_summary_spring, iRts, spring_last_date, cumIJan, k_tr_adv_distrib, k_p_distrib, c_distrib, p_curr_distrib)



df_t_sim2<-out_list[[1]] # contains the time course of runs for 500 simulations for both scenarios
df_t2<-out_list[[2]] # contains the median and bounds of the time course runs for both scenarios
df_summary<-out_list[[3]] # contains all the summary statistics for static variables
df_t_sim['I_variant_lit']<-df_t_sim2$I_variant
df_t['Imedian_lit']<-df_t2$Imedianv

```
Troubleshoot davies et al
```{r}
springdates <- seq(Rt_summary_spring$dates[1], spring_end_semester_date, "days")
 tsim<-seq(from = 0, to = length(springdates)-1, by = dt)
  f_ts<-matrix(nrow = length(tsim), ncol = nsamps)
  p_ts<-matrix(nrow = length(tsim), ncol = nsamps)
  
for( i in 1:nsamps){
    p_curr<-p_curr_distrib[i] 
    k<-kl_distrib[i]
    c<-c2_distrib[i]
    tr_adv<-exp(mean_GI*k) -1 # M-1
    
    #tr_adv<-tr_adv_distrib[i] function of  exp(b1*g_time)
    # f_t must correspond to the correct b0, b1, pcurr, and tr_adv parameters
    f_ts[,i]<-get_ft(c, k, p_curr, tr_adv, tsim)
    p_ts[,i] = (1/(1+exp(-(c+k*tsim))))
    
  }
  
  
  f_t_median<-apply( f_ts , 1 , quantile , probs = 0.5 , na.rm = TRUE )
  f_t_lb<-apply( f_ts , 1 , quantile , probs = 0.025 , na.rm = TRUE )
  f_t_ub<-apply( f_ts , 1 , quantile , probs = 0.975 , na.rm = TRUE )
  df_test<-data.frame(tsim, f_t_median, f_t_lb, f_t_ub)
  
test<-ggplot(df_test) + geom_line(aes(x = tsim, y = f_t_median)) + 
  geom_ribbon(aes(x =tsim, ymin = f_t_lb, ymax = f_t_ub), alpha = 0.1)
test
```
Make some plots of the infection projection outputs
```{r}
#Factor increase in transmission rate over time
path = "../out/factor_inc_t.png"
plt_ft<-ggplot() + geom_line(data = df_t, aes (x = springdates, y = f_t_median)) + geom_ribbon(data = df_t, aes(x = springdates, ymin = f_t_lb, ymax = f_t_ub), alpha = 0.1) +geom_line(data = df_t2, aes (x = springdates, y = f_t_median), color = 'red') + geom_ribbon(data = df_t2, aes(x = springdates, ymin = f_t_lb, ymax = f_t_ub), alpha = 0.1, fill = "red")+
  geom_vline(xintercept = as.Date(spring_last_date))+ xlab('Time') + ylab('Multiplicative factor on transmission rate')
plt_ft

N<-N_POPULATION
case_data_spring['calib']<-rep('fit', nrow(case_data_spring))
for(i in 1:nrow(case_data_spring)){
  if (case_data_spring$date[i]>spring_last_date){
    case_data_spring$calib[i]<-'observed'
  }
}

ggsave(filename = path, device = "png", dpi = 700)

# Infection estimates alongside observed test positivity
case_data_spring<-case_data_spring%>%filter(date<=spring_last_date)
path = "../out/both_scenarios.png"
plot1<-ggplot() + geom_line(data = df_t_sim, aes(x = dates, y = 1000*I_variant/N, group = samples), alpha = 0.05, color = "tan1") + geom_line(data = df_t_sim, aes(x = dates, y = 1000*I_variant_lit/N, group = samples), alpha = 0.03, color = "darkolivegreen4")+
  geom_line(data = df_t_sim, aes(x = dates, y = 1000*I_novariant/N, group = samples), alpha = 0.03, color = "mediumpurple") +
  geom_line(data = df_t, aes(x = springdates, y = 1000*Imediannv/N), size = 1, color = "mediumpurple")+ geom_line(data = df_t, aes(x = springdates, y = 1000*Imedianv/N), size = 1, color = "tan1") +geom_line(data = df_t, aes(x = springdates, y = 1000*Imedian_lit/N), size = 1, color = "darkolivegreen4")+
  geom_point(data = case_data_spring, aes(x = date, y = 1000*sevendayPCTpos), size = 0.5) +facet_wrap(~scenario) + 
  xlab ("Date") + ylab('COVID-19 infections per 1000')+theme(strip.background = element_blank(), aspect.ratio = 0.8) + coord_cartesian(ylim = c(0, 50))
plot1
ggsave(filename = path, device = "png", dpi = 700)

df_fast<-df_t_sim%>%filter(scenario=="Faster spread")
df_t_fast<-df_t%>%filter(scenario == "Faster spread")
path ="../out/fast_spaghettis.png"
fast_only<-ggplot() + geom_line(data = df_fast, aes(x = dates, y = 1000*I_variant/N, group = samples), alpha = 0.05, color = "tan1") +
  geom_line(data = df_fast, aes(x = dates, y = 1000*I_novariant/N, group = samples), alpha = 0.03, color = "mediumpurple") +
   geom_line(data = df_fast, aes(x = dates, y = 1000*I_variant_lit/N, group = samples), alpha = 0.03, color = "darkolivegreen4")+
  geom_line(data = df_t_fast, aes(x = springdates, y = 1000*Imediannv/N), size = 1, color = "mediumpurple")+ geom_line(data = df_t_fast, aes(x = springdates, y = 1000*Imedianv/N), size = 1, color = "tan1") +geom_line(data = df_t_fast, aes(x = springdates, y = 1000*Imedian_lit/N), size = 1, color = "darkolivegreen4")+
  geom_point(data = case_data_spring, aes(x = date, y = 1000*sevendayPCTpos), size = 1) + scale_x_date(date_labels="%b %d")+
  xlab ("Date") + ylab("") +theme(strip.background = element_blank())+ coord_cartesian(ylim = c(0, 50)) + ylab('COVID-19 infections per 1000') + xlab ("Date")
fast_only
ggsave(filename = path, device = "png", dpi = 1000, width = 7, height = 5, units="in")
save.image(file="foobar.Rdata")


# all 4 scenarios median model lines
path = "../out/all4medians.png"
plot4<-ggplot() + geom_line(data = df_t, aes(x = springdates, y = 1000*Imediannv/N, group = scenario, linetype = scenario),color = "mediumpurple", size = 1) + 
  geom_line(data = df_t, aes(x = springdates, y = 1000*Imedianv/N, group = scenario, linetype = scenario), color = "tan1", size = 1) +geom_line(data = df_t, aes(x = springdates, y = 1000*Imedian_lit/N, group = scenario, linetype = scenario), color = "darkolivegreen4", size = 1)+
  xlab ("Date") + ylab('COVID-19 infections per 1000')+theme(aspect.ratio = 0.8) + coord_cartesian(ylim = c(0, 50))+ ggtitle("Median model lines by scenario")
plot4
ggsave(filename = path, device = "png", dpi = 700)



# imputed infections B117 and WT
path = "../out/both_inf_t.png"
plot_both<-ggplot() + geom_line(data = df_t, aes(x=springdates, y = 1000*B117medianv/N), color = "red") + 
  geom_ribbon(data = df_t, aes(x = springdates, ymin = 1000*B117lbv/N, ymax = 1000*B117ubv/N), alpha = 0.1, fill = "red") +
  geom_line(data = df_t, aes(x=springdates, y = 1000*WTmedianv/N), color = "blue")+
  geom_ribbon(data = df_t, aes(x = springdates, ymin = 1000*WTlbv/N, ymax = 1000*WTubv/N), alpha = 0.1, fill = "blue")+
  geom_line(data = df_t, aes(x=springdates, y = 1000*Imedianv/N), color = "tan1")+
  geom_ribbon(data = df_t, aes(x = springdates, ymin = 1000*Ilbv/N, ymax = 1000*Iubv/N), alpha = 0.1, fill = "tan1")+
  facet_wrap(~scenario) + xlab ("Date") + ylab('COVID-19 infections per 1000')  + scale_y_log10() + coord_fixed(ratio = 50) + theme(axis.title.y = element_text(size = 10))
plot_both
ggsave(filename = path, device = "png", dpi = 700)

plot_both_lit<- ggplot() + geom_line(data = df_t2, aes(x=springdates, y = 1000*B117medianv/N), color = "red") + 
  geom_ribbon(data = df_t2, aes(x = springdates, ymin = 1000*B117lbv/N, ymax = 1000*B117ubv/N), alpha = 0.1, fill = "red") +
  geom_line(data = df_t2, aes(x=springdates, y = 1000*WTmedianv/N), color = "blue")+
  geom_ribbon(data = df_t2, aes(x = springdates, ymin = 1000*WTlbv/N, ymax = 1000*WTubv/N), alpha = 0.1, fill = "blue")+
  geom_line(data = df_t2, aes(x=springdates, y = 1000*Imedianv/N), color = "darkolivegreen4")+
  geom_ribbon(data = df_t2, aes(x = springdates, ymin = 1000*Ilbv/N, ymax = 1000*Iubv/N), alpha = 0.1, fill = "darkolivegreen4")+ scale_y_log10() + 
  facet_wrap(~scenario) + xlab ("Date") + ylab('COVID-19 infections per 1000') + coord_fixed(ratio = 50) + theme(axis.title.y = element_text(size = 10))
plot_both_lit 
path = "../out/B117_WT_both.png"
figure<-ggarrange(plot_both, plot_both_lit, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right")
figure
ggsave(filename = path, device = "png", dpi = 700)
```



Make figures with frequency of SGTF over time as well as cases colored by SGTF status
```{r}
# Make figures with frequency of SGTF over time as well as cases colored by SGTF status
path ="../out/dailySGTF.png"
plt_freq<-ggplot(Ct_data_daily_summary, aes(x = test_date, y = 100*(prev_SGTF)), color = "black") + geom_point() + geom_errorbar(aes(ymin = 100*(prev_SGTF - SE_prev), ymax = 100*(prev_SGTF+SE_prev)), width = 0.2, position = position_dodge(0.5))+
xlab ('Test Collection Date') + ylab ('Frequency (%) of cases that are S dropouts') + ggtitle('Daily frequency of S dropouts with low Ct values')
plt_freq
ggsave(filename = path, device = "png", dpi = 700) 

path = "../out/stackedbarSGTF.png"
stacked<-ggplot(Ct_data_daily_long, aes(x = test_date, y = n, fill = strain))  + geom_bar(position = "stack", stat = "identity") + xlab('Test Collection Date') + ylab('Number of low Ct positives') + theme(legend.position="top")+ scale_fill_discrete(name = "S gene presence", labels = c("S+", "S-"))
stacked
ggsave(filename = path, device = "png", dpi = 700)
```










Exploratory data analysis of the SGTF vs non-SGTF viral loads
```{r}
# Make a violin plot comparing the Ct values of side-by-side SGTF to non-SGTF
path="../out/Ct_distribs.png"
Ct_values<-ggplot(data=Ct_data, aes(factor(strain), mean_Ct), group = strain)+ geom_violin(position = position_dodge(width = 0.9)) + geom_boxplot(width = 0.1)+geom_jitter(shape = 16, alpha = 0.3, position = position_jitter(width=NULL, height = NULL))+ geom_hline(yintercept = QC_thres, linetype = "dashed") + xlab('Possible Strain') + ylab('Ct value') + ggtitle('Ct values by strain') 
Ct_values
ggsave(filename = path, device = "png", dpi = 700)

# Summary statistics
n_SGTF<-as.numeric(sum(Ct_data%>%filter(S_dropout == TRUE)%>%summarise(n())))
n_non_SGTF<-nrow(Ct_data)-n_SGTF
n_SGTF_low_Ct<-as.numeric(sum(Ct_data%>%filter(S_dropout==TRUE & mean_Ct<QC_thres)%>%summarise(n())))
n_non_SGTF_low_Ct<-as.numeric(sum(Ct_data%>%filter(S_dropout==FALSE & mean_Ct<QC_thres)%>%summarise(n())))
Ct_all_combined<-Ct_data%>%group_by(strain)%>%
  summarise(median_Ct = quantile(mean_Ct, 0.5),
    Ct_lb = quantile(mean_Ct, 0.025),
    Ct_ub = quantile(mean_Ct, 0.975),
    n_samps = n())

# Run a Kolmogorov-Smirnov test to see if there is a statistically significant difference in the Ct value distributions
ks.test(Ct_data$mean_Ct[Ct_data$strain == "SGTF"], Ct_data$mean_Ct[Ct_data$strain == "non-SGTF"])
```