---
title: "UT_B117_main"
output: html_document
---
This script loads in the raw data files and calls the functions that run the logistic growth model of B117 vs wildtype and the SEIR model the projects infections
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(EpiEstim)
library(tidyr)
library(reshape2)
library(cowplot)
library(knitr)
library(flexmix)
library(zoo)
library(DescTools)
library(dplyr)
source("UT_B117_functions.R")
theme_set(theme_cowplot())
ggplot + background_grid(major = "xy", minor = "xy")
options(warn = -1)
```
Load in SGTF data and PCT case data
```{r}
DATA_PATH_CTS<-'../data/pctData_posOnly_210505.csv'
DATA_PATH_CASES<-'../data/UT_data_streams.csv'
DATA_PATH_INTROS<-'../data/introductions_Jan15.csv'

```
Dates and parameters for analysis
```{r}
spring_arrival_date<-as.Date("2021-01-16", format = "%Y-%m-%d")
spring_last_date<-as.Date("2021-04-09", format = "%Y-%m-%d")
spring_end_semester_date<-as.Date("2021-05-23", format = "%Y-%m-%d")
initial_estimate_date<-as.Date("2021-02-12", format = "%Y-%m-%d")
fall_arrival_date<-as.Date("2020-08-20", format = "%Y-%m-%d")
fall_last_date<-as.Date("2020-12-19", format = "%Y-%m-%d")

# Some global parameters
QC_thres<-28
delta<-1/7 #(1/duration of infectiousness)
gamma<- 1/3 # (1/duration of latent period)
dt<-0.1
mean_GI<-5.8 # average generation interval (for EpiEstim)
std_GI<-2.3 # std dev generation interval (for EpiEstim)
N_POPULATION <-30000 # estimated "active" population at UT
prop_prev_inf_Aug = 0.05 
nsamps<-500 # number of samples from the distributions for uncertainty
```
Process the Ct data to get SGTF and non-SGTF positives from the Ct values of the 3 target genes
```{r}
Ct_data<-read.csv(DATA_PATH_CTS) %>%
  mutate(test_date = as.Date(test_date))
Ct_data<-Ct_data %>%select(test_date, N.gene.Ct, S.gene.Ct, ORF1ab.Ct)%>%
  filter(test_date>= spring_arrival_date)%>%
  filter(test_date<=spring_last_date)
Ct_data<-Ct_data[order(Ct_data$test_date),]
Ct_data<-Ct_data%>%rowwise()%>%mutate(mean_Ct = mean(c(N.gene.Ct, S.gene.Ct, ORF1ab.Ct), na.rm = TRUE))
Ct_data['S_dropout']<-is.na(Ct_data$S.gene.Ct)
Ct_data$poss_B117<-Ct_data$S_dropout==TRUE & Ct_data$N.gene.Ct<QC_thres
Ct_data['strain']<-rep(NA, nrow(Ct_data))
for (i in 1:nrow(Ct_data)){
  if (Ct_data$S_dropout[i] == TRUE){
    Ct_data$strain[i]<-'SGTF'
  }
  else 
    Ct_data$strain[i]<-'non_SGTF'
}

#Make a day to day summary of SGTF presence
Ct_data_daily_long<-Ct_data%>%filter(mean_Ct<QC_thres)%>%
  group_by(test_date, strain)%>%summarise(n = n())
Ct_data_daily_summary<-Ct_data_daily_long%>%
  pivot_wider(names_from = strain, values_from = n, values_fill = 0)
Ct_data_daily_summary<-Ct_data_daily_summary%>%group_by(test_date)%>%
  mutate(n_Sdrop = SGTF,
         n_Spos = non_SGTF,
         n_pos = SGTF+non_SGTF,
         prev_SGTF = SGTF/n_pos,
         SE_prev = sqrt(prev_SGTF*(1-prev_SGTF)/n_pos))
# Export the daily data for fitting. Will use this to feed into Woody's SGTF.R code
write.csv(Ct_data_daily_summary, '../data/Ct_data.csv')
# Make a dataset used for the initial fitting from data through Feb.
Ct_data_init<-Ct_data_daily_summary%>%filter(test_date<=initial_estimate_date)
write.csv(Ct_data_init, '../data/Ct_data_init.csv', row.names = FALSE)


# Make figures with frequency of SGTF over time as well as cases colored by SGTF status
path ="../out/dailySGTF.png"
plt_freq<-ggplot(Ct_data_daily_summary, aes(x = test_date, y = 100*(prev_SGTF)), color = "black") + geom_point() + geom_errorbar(aes(ymin = 100*(prev_SGTF - SE_prev), ymax = 100*(prev_SGTF+SE_prev)), width = 0.2, position = position_dodge(0.5))+
xlab ('Test Collection Date') + ylab ('Frequency (%) of cases that are S dropouts') + ggtitle('Daily frequency of S dropouts with low Ct values')
plt_freq
ggsave(filename = path, device = "png", dpi = 700) 

path = "../out/stackedbarSGTF.png"
stacked<-ggplot(Ct_data_daily_long, aes(x = test_date, y = n, fill = strain))  + geom_bar(position = "stack", stat = "identity") + xlab('Test Collection Date') + ylab('Number of low Ct positives') + theme(legend.position="top")+ scale_fill_discrete(name = "S gene presence", labels = c("S+", "S-"))
stacked
ggsave(filename = path, device = "png", dpi = 700)

# Use SGTF data set to call function to fit the logistic model with uncertainty from SGTF presence. 

#output: c and k in logistic growth model (intercept and growth rate), current B117 prevalence
c = -2.861
c_distrib<-runif(n= nsamps, min = -2.9, max = -2.83)
k = 0.040
k_distrib = runif(n=nsamps, min = 0.037, max = 0.043)
p_curr = 0.627
p_curr_distrib<-runif(nsamps, 0.501, 0.739)
M_distrib = exp(mean_GI*k_distrib) # This should probably be a function of b0 and b1? 


```
Process the daily case data from UT students
```{r}
case_data<-read.csv(DATA_PATH_CASES)%>%mutate(date = as.Date(CollectionDate, format = "%Y-%m-%d"))%>%
                                                select(date, student_pos, student_PCT_pos,
                                                       total_student_PCT_tests)%>%
                                                filter(date>=fall_arrival_date, date<=spring_last_date+7)
# Add some columns to the case data
case_data<-case_data%>%mutate(positivity = student_PCT_pos/total_student_PCT_tests)
sevendayPCTpos = c(rep(NA,3), rollapply(case_data$positivity, 7, function(x) mean(x, na.rm = TRUE)), rep(NA,3))
sevendayPCTcases = c(rep(NA,3), rollapply(case_data$student_PCT_pos, 7, function(x) mean(x, na.rm = TRUE)), rep(NA,3))
sevendaycases = c(rep(NA,3), rollapply(case_data$student_pos, 7, function(x) mean(x, na.rm = TRUE)), rep(NA,3))
case_data$sevendayPCTpos = sevendayPCTpos
case_data$sevendayPCTcases = sevendayPCTcases
case_data$sevendaycases = sevendaycases

case_data%>%ggplot() + geom_line(aes(x = date, y = sevendayPCTpos)) + geom_bar(aes(x = date, y = positivity), stat = "identity") + ylab('PCT positivity') + xlab('Test Collection Date')

case_data%>%ggplot() + geom_line(aes(x = date, y = sevendayPCTcases)) + geom_bar(aes(x = date, y = student_PCT_pos), stat = "identity") + ylab('Student PCT cases') + xlab('Test Collection Date')

# Estimate R(t) in the fall and in the spring.
#daily_p_local<-get_daily_p_local(case_data, fall_arrival_date)
case_data_fall<-case_data%>%filter(date<=fall_last_date)
case_data_spring<-case_data%>%filter(date>=spring_arrival_date)
cases_fall<-case_data_fall$student_pos # for fall, we use total positives

avg_tests_per_day<-mean(case_data_spring$total_student_PCT_tests, na.rm = TRUE)
case_data_spring$positivity[is.na(case_data_spring$positivity)]=case_data_spring$sevendayPCTpos[is.na(case_data_spring$positivity)]
# Still need to do something to account for 10-day pause during the Texas winter freeze where no tests were collected, use the 3 days before and after it to impute. 
case_data_around_freeze<-case_data%>%filter(date<=as.Date("2021-02-24"), date>=as.Date("2021-02-10"),
                                            total_student_PCT_tests>0)
avg_positivity_freeze<-sum(case_data_around_freeze$student_PCT_pos)/sum(case_data_around_freeze$total_student_PCT_tests)
avg_n_cases_freeze<-sum(case_data_around_freeze$student_pos)/nrow(case_data_around_freeze)
case_data_spring$positivity[is.na(case_data_spring$positivity)]=avg_positivity_freeze
cases_spring<-round(avg_tests_per_day*case_data_spring$positivity,0) # for spring, we used PCT positivity
case_data_spring$imputed_cases<-cases_spring


Rt_summary_fall<-Rt_fxn_cases(cases_fall, case_data_fall, fall_arrival_date, fall_last_date, mean_GI,std_GI)
Rt_summary_spring<-Rt_fxn_cases(cases_spring, case_data_spring, spring_arrival_date, spring_last_date+7, mean_GI,std_GI)

# Export the Rt data for use in the SEIR model
write.csv(Rt_summary_fall, '../data/Rt_students_fall.csv', row.names = FALSE)
write.csv(Rt_summary_spring, '../data/Rt_students_spring.csv', row.names = FALSE)

Rt_summary_fall%>%ggplot() + geom_line(aes(x =dates, y = Rt_medians))+ geom_ribbon(aes(x = dates, ymin = Rt_lowerbounds, ymax = Rt_upperbounds), alpha = 0.1)
Rt_summary_spring%>%ggplot() + geom_line(aes(x =dates, y = Rt_medians))+ geom_ribbon(aes(x = dates, ymin = Rt_lowerbounds, ymax = Rt_upperbounds), alpha = 0.1)
```
```{r}
path = "../out/student_cases.png"
geom_mean<- exp(rollapply(log(case_data_spring$imputed_cases/N_POPULATION), 7, mean,na.rm = TRUE))
geom_mean<-c(rep(NA, 3), geom_mean, rep(NA, 3))
case_data_spring$geom_mean<-geom_mean
cases<-ggplot(case_data_spring) + geom_line(aes(x = date, y = sevendaycases)) + geom_bar(aes(x = date, y = student_pos), stat = "identity") + xlab('Date') + ylab('Total student cases') + scale_y_continuous(trans='log2')
cases
ggsave(filename = path, device = "png", dpi = 700)

tspring<-seq(from = 0, to = nrow(case_data_spring)-1, by =1)
 p_t<-(1/(1+exp(-(c+k*tspring))))
plot(tspring, p_t)

path = "../out/student_cases_by_B117.png"
case_data_spring$n_est_B117<-case_data_spring$student_pos*p_t
case_data_spring$n_est_WT<-case_data_spring$student_pos*(1-p_t)
case_data_long<-case_data_spring%>%select(date, n_est_B117, n_est_WT)%>%
  pivot_longer(cols = starts_with("n_est_"), names_to = "strain", names_prefix = "n_est_", values_to = "n", values_drop_na = TRUE)
cases<-ggplot(case_data_long, aes(x = date, y = n, fill = strain)) + geom_bar(position = "stack", stat = "identity") + xlab('Date') + ylab('Total student cases') 
cases
ggsave(filename = path, device = "png", dpi = 700)

```
Run the SEIR model using the fall case data to estimate the number previously infected
```{r}
fall_intros<-read.csv(DATA_PATH_INTROS)%>%filter(Location == "total", Time == "Fall")%>%select(ML_infections_lo, ML_infections_med, ML_infections_hi)
I0bounds<-c(fall_intros$ML_infections_lo, fall_intros$ML_infections_hi)

par_table<-data.frame(dt, delta, gamma, N_POPULATION, prop_prev_inf_Aug, nsamps)

cumIJan<-run_fall_SEIR(par_table, I0bounds, Rt_summary_fall)
cumICI<-(cumIJan/N_POPULATION)%>%quantile(probs = c(0.025, 0.5, 0.975))
print(cumICI)
```
Use the previous infection estimates from the fall, spring Rt, the current estimates of variant prevalence, and variant increase in transmissibility to project infections with and without the variant.
```{r}

spring_par_table<-data.frame(dt, delta, gamma, N_POPULATION, nsamps)
p_curr_distrib<-runif(nsamps, 0.501, 0.739) # This should be from Woody's output, this is a placeholder
iRts<-c(18, 14) # indices in spring Rt that correspond to the faster and slower spread scenarios

out_list<-run_spring_SEIR(spring_par_table, case_data_spring, Rt_summary_spring, iRts, spring_last_date, cumIJan, k_distrib,c_distrib, p_curr_distrib)

df_t_sim<-out_list[[1]] # contains the time course of runs for 500 simulations for both scenarios
df_t<-out_list[[2]] # contains the median and bounds of the time course runs for both scenarios
df_summary<-out_list[[3]] # contains all the summary statistics for static variables

```
Make some plots of the infection projection outputs
```{r}
#Factor increase in transmission rate over time
path = "../out/factor_inc_t.png"
plt_ft<-ggplot(df_t) + geom_line(aes (x = tsim, y = f_t_median)) + geom_ribbon(aes(x = tsim, ymin = f_t_lb, ymax = f_t_ub), alpha = 0.1) + xlab('Time') + ylab('Multiplicative factor on transmission rate')
plt_ft

N<-N_POPULATION
case_data_spring['calib']<-rep('fit', nrow(case_data_spring))
for(i in 1:nrow(case_data_spring)){
  if (case_data_spring$date[i]>spring_last_date){
    case_data_spring$calib[i]<-'observed'
  }
}

ggsave(filename = path, device = "png", dpi = 700)

# Infection estimates alongside observed test positivity
case_data_spring<-case_data_spring%>%filter(date<=spring_last_date)
path = "../out/both_scenarios.png"
plot1<-ggplot() + geom_line(data = df_t_sim, aes(x = dates, y = 1000*I_variant/N, group = samples), alpha = 0.05, color = "tan1") +
  geom_line(data = df_t_sim, aes(x = dates, y = 1000*I_novariant/N, group = samples), alpha = 0.02, color = "mediumpurple") +
  geom_line(data = df_t, aes(x = springdates, y = 1000*Imediannv/N), size = 1, color = "mediumpurple")+ geom_line(data = df_t, aes(x = springdates, y = 1000*Imedianv/N), size = 1, color = "tan1") +
  geom_point(data = case_data_spring, aes(x = date, y = 1000*sevendayPCTpos), size = 0.5) +facet_wrap(~scenario) + 
  xlab ("Date") + ylab('COVID-19 infections per 1000')+theme(strip.background = element_blank(), aspect.ratio = 0.8) + coord_cartesian(ylim = c(0, 50))
plot1
ggsave(filename = path, device = "png", dpi = 700)


# all 4 scenarios median model lines
path = "../out/all4medians.png"
plot4<-ggplot() + geom_line(data = df_t, aes(x = springdates, y = 1000*Imediannv/N, group = scenario, linetype = scenario),color = "tan1", size = 1) + 
  geom_line(data = df_t, aes(x = springdates, y = 1000*Imedianv/N, group = scenario, linetype = scenario), color = "mediumpurple", size = 1) +
  xlab ("Date") + ylab('COVID-19 infections per 1000')+theme(aspect.ratio = 0.8) + coord_cartesian(ylim = c(0, 50))+ ggtitle("Median model lines by scenario")
plot4
ggsave(filename = path, device = "png", dpi = 700)


# imputed infections B117 and WT
path = "../out/both_inf_t.png"
plot_both<-ggplot() + geom_line(data = df_t, aes(x=springdates, y = B117medianv), color = "red") + 
  geom_ribbon(data = df_t, aes(x = springdates, ymin = B117lbv, ymax = B117ubv), alpha = 0.1, fill = "red") +
  geom_line(data = df_t, aes(x=springdates, y = WTmedianv), color = "blue")+
  geom_ribbon(data = df_t, aes(x = springdates, ymin = WTlbv, ymax = WTubv), alpha = 0.1, fill = "blue")+
  geom_line(data = df_t, aes(x=springdates, y = Imedianv), color = "black")+
  geom_ribbon(data = df_t, aes(x = springdates, ymin = Ilbv, ymax = Iubv), alpha = 0.1, fill = "black")+
  facet_wrap(~scenario) + xlab ("Date") + ylab('COVID-19 infections')
plot_both
ggsave(filename = path, device = "png", dpi = 700)

path = "../out/both_inf_log.png"
plot_bothlog<-ggplot() + geom_line(data = df_t, aes(x=springdates, y = B117medianv), color = "red") + 
  geom_ribbon(data = df_t, aes(x = springdates, ymin = B117lbv, ymax = B117ubv), alpha = 0.1, fill = "red") +
  geom_line(data = df_t, aes(x=springdates, y = WTmedianv), color = "blue")+
  geom_ribbon(data = df_t, aes(x = springdates, ymin = WTlbv, ymax = WTubv), alpha = 0.1, fill = "blue")+
  geom_line(data = df_t, aes(x=springdates, y = Imedianv), color = "black")+
  geom_ribbon(data = df_t, aes(x = springdates, ymin = Ilbv, ymax = Iubv), alpha = 0.1, fill = "black")+
  facet_wrap(~scenario) + xlab ("Date") + ylab('COVID-19 infections') + scale_y_log10()
plot_bothlog
ggsave(filename = path, device = "png", dpi = 700)
```


# Only faster spread scenario
df_fast<-df_t_sim%>%filter(scenario=="Faster spread")
df_t_fast<-df_t%>%filter(scenario == "Faster spread")
path ="../out/fast_spaghettis.png"
fast_only<-ggplot() + geom_line(data = df_fast, aes(x = dates, y = 1000*I_variant/N, group = samples), alpha = 0.05, color = "tan1") +
  geom_line(data = df_fast, aes(x = dates, y = 1000*I_novariant/N, group = samples), alpha = 0.02, color = "mediumpurple") +
  geom_line(data = df_t_fast, aes(x = springdates, y = 1000*Imediannv/N), size = 1, color = "mediumpurple")+ geom_line(data = df_t_fast, aes(x = springdates, y = 1000*Imedianv/N), size = 1, color = "tan1") +
  geom_point(data = case_data_spring, aes(x = date, y = 1000*sevendayPCTpos), size = 1) + scale_x_date(date_labels="%b-%d")+
  xlab ("Date") + ylab("") +theme(strip.background = element_blank())+ coord_cartesian(ylim = c(0, 50)) + ylab('COVID-19 infections per 1000') + xlab ("Date")
fast_only
ggsave(filename = path, device = "png", dpi = 1000, width = 7, height = 5, units="in")
save.image(file="foobar.Rdata")
```










Make figures with frequency of SGTF over time as well as cases colored by SGTF status
```{r}
# Make figures with frequency of SGTF over time as well as cases colored by SGTF status
path ="../out/dailySGTF.png"
plt_freq<-ggplot(Ct_data_daily_summary, aes(x = test_date, y = 100*(prev_SGTF)), color = "black") + geom_point() + geom_errorbar(aes(ymin = 100*(prev_SGTF - SE_prev), ymax = 100*(prev_SGTF+SE_prev)), width = 0.2, position = position_dodge(0.5))+
xlab ('Test Collection Date') + ylab ('Frequency (%) of cases that are S dropouts') + ggtitle('Daily frequency of S dropouts with low Ct values')
plt_freq
ggsave(filename = path, device = "png", dpi = 700) 

path = "../out/stackedbarSGTF.png"
stacked<-ggplot(Ct_data_daily_long, aes(x = test_date, y = n, fill = strain))  + geom_bar(position = "stack", stat = "identity") + xlab('Test Collection Date') + ylab('Number of low Ct positives') + theme(legend.position="top")+ scale_fill_discrete(name = "S gene presence", labels = c("S+", "S-"))
stacked
ggsave(filename = path, device = "png", dpi = 700)
```










Exploratory data analysis of the SGTF vs non-SGTF viral loads
```{r}
# Make a violin plot comparing the Ct values of side-by-side SGTF to non-SGTF
path="../out/Ct_distribs.png"
Ct_values<-ggplot(data=Ct_data, aes(factor(strain), mean_Ct), group = strain)+ geom_violin(position = position_dodge(width = 0.9)) + geom_boxplot(width = 0.1)+geom_jitter(shape = 16, alpha = 0.3, position = position_jitter(width=NULL, height = NULL))+ geom_hline(yintercept = QC_thres, linetype = "dashed") + xlab('Possible Strain') + ylab('Ct value') + ggtitle('Ct values by strain') 
Ct_values
ggsave(filename = path, device = "png", dpi = 700)

# Summary statistics
n_SGTF<-as.numeric(sum(Ct_data%>%filter(S_dropout == TRUE)%>%summarise(n())))
n_non_SGTF<-nrow(Ct_data)-n_SGTF
n_SGTF_low_Ct<-as.numeric(sum(Ct_data%>%filter(S_dropout==TRUE & mean_Ct<QC_thres)%>%summarise(n())))
n_non_SGTF_low_Ct<-as.numeric(sum(Ct_data%>%filter(S_dropout==FALSE & mean_Ct<QC_thres)%>%summarise(n())))
Ct_all_combined<-Ct_data%>%group_by(strain)%>%
  summarise(median_Ct = quantile(mean_Ct, 0.5),
    Ct_lb = quantile(mean_Ct, 0.025),
    Ct_ub = quantile(mean_Ct, 0.975),
    n_samps = n())

# Run a Kolmogorov-Smirnov test to see if there is a statistically significant difference in the Ct value distributions
ks.test(Ct_data$mean_Ct[Ct_data$strain == "SGTF"], Ct_data$mean_Ct[Ct_data$strain == "non-SGTF"])
```